---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r Loading and preprocessing data, echo=TRUE,results='hide'}
#Libraries to help with data preprocessing and plotting
library(plyr)
library(dplyr)
library(lubridate)
library(ggplot2)

activity <- read.csv("~/R/RepData_PeerAssessment1/activity.csv")

#Change date from factor to POSIXct
activity<- activity %>% mutate(date=ymd(activity$date))
```


## What is mean total number of steps taken per day?
```{r Mean steps per day, echo=TRUE, results='hide'}
#Data summarized to report sum, mean and median steps per day
steps_per_day<- na.omit(activity) %>% group_by(date) %>% summarize(sum(steps), mean(steps), median(steps))
names(steps_per_day)<-c("date","total_steps","mean","median")
```

```{r echo=TRUE}
mean(steps_per_day$total_steps)
```

```{r echo=TRUE}
median(steps_per_day$steps)
```

```{r histogram_steps_per_day, fig.height=4,fig.width=6}
ggplot(steps_per_day,aes(x=total_steps))+ geom_histogram(aes(fill = ..count..)) + scale_fill_gradient("Count", low = "green", high = "red")
```

## What is the average daily activity pattern?
```{r steps per interval, echo=TRUE}
steps_interval<-activity%>%group_by(interval)%>%summarize(mean(steps,na.rm=TRUE))
names(steps_interval)<-c("interval","average_steps")
```

```{r plot average steps per interval, fig.height=4,fig.width=6}
ggplot(data=steps_interval,aes(x=interval,y=average_steps)) + geom_line(color="blue")
```

Interval with maximum number of average steps
```{r}
steps_interval[which.max(steps_interval$average_steps),]$interval
```

## Imputing missing values

Total missing values
```{r echo=TRUE}
sum(is.na(activity$steps))
```

Missing values imputed by substituting for average steps for interval
```{r echo=TRUE}
steps_for_interval<-function(interval){
    steps_interval[steps_interval$interval==interval,]$steps
}
```

New data set with missing values filled
```{r echo=TRUE}
new_activity<-activity
count=0
for(i in 1:nrow(new_activity)){
    if(is.na(new_activity[i,]$steps)){
        new_activity[i,]$steps<-steps_for_interval(new_activity[i,]$interval)
        count=count+1
    }
}
cat("Total ",count, "NA values were filled")  
```

Histogram of new activity data
```{r Mean steps per day2, echo=TRUE, results='hide'}
#Data summarized to report sum, mean and median steps per day
steps_per_day2<- new_activity %>% group_by(date) %>% summarize(sum(steps), mean(steps), median(steps))
names(steps_per_day2)<-c("date","total_steps","mean","median")
```

```{r echo=TRUE}
mean(steps_per_day2$total_steps)
```

```{r echo=TRUE}
median(steps_per_day2$steps)
```

```{r histogram_steps_per_day2, fig.height=4,fig.width=6}
ggplot(steps_per_day2,aes(x=total_steps)) + geom_histogram(aes(fill = ..count..)) + scale_fill_gradient("Count", low = "green", high = "red")
```

## Are there differences in activity patterns between weekdays and weekends?
